---
title: "p8105_hw1_kx2224"
author: "Kangyu Xu (kx2224)"
date: "2024-09-18"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
## Question 1
### Load the data
```{r}
data("penguins", package = "palmerpenguins") 
```
### Short Description
The `penguins` dataset contains information about different species of penguins, including variables such as `species`, `island`, `bill_length_mm`, `bill_depth_mm`, `flipper_length_mm`, `body_mass_g`, `sex` and `year`. The dataset consists of `r nrow(penguins)` rows and `r ncol(penguins)` columns. The mean flipper length of the penguins is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm.

### Make a scatterplot
```{r}
library(ggplot2)
firstScatterPlot = ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, colour = species))+
  geom_point(na.rm = TRUE)+
  labs(title = "Scatter Plot of Flipper_length_mm (y) vs Bill_length_mm (x)")
```

### Save the plot
```{r}
ggsave("Scatter Plot of Flipper_length_mm (y) vs Bill_length_mm (x).jpg", plot = firstScatterPlot)
```

## Question 2
### Create the dataframe
```{r}
library(tidyverse)
set.seed(42)
randomSample = rnorm(10)
logicVector = randomSample>0
charVector = sample(letters, 10, replace = TRUE)
factorVector = sample(c("level1","level2","level3"), 10, replace = TRUE)
df = data.frame(randomSample, logicVector, charVector, factorVector)
df
```
### Calculate the mean of variables
```{r}
mean_randomSample = mean(df %>% pull(randomSample))
mean_logicVector = mean(df %>% pull(logicVector))
mean_charVector = try(mean(df %>% pull(charVector)), silent = TRUE) 
mean_factorVector = try(mean(df %>% pull(factorVector)), silent = TRUE) 

# output the result
print(mean_randomSample)
print(mean_logicVector)
print(mean_charVector)
print(mean_factorVector)
```
In conclusion, we can take the mean of "numeric" and "logical vector". But for "character" and "factor", the R cannot calculate the mean of them.

### Convert variables
```{r}
numeric_logical = as.numeric(logicVector)
numeric_character = try(as.numeric(charVector), silent = TRUE)
numeric_factor = try(as.numeric(factorVector), silent = TRUE)

# output the result
print(numeric_logical)
print(numeric_character)
print(numeric_factor)
```
The result of conversion is that we succeeded to convert logical vector, but we failed to convert both factors and charactors. The reason is that logical values are coerced to 1 (TRUE) and 0 (FALSE), so this conversion works. But for characters and factors, they cannot be directly converted to numerics, so the result will be NA as showed above.

This result can help explain why we could only take the mean of logical vectors. 